---
title: "Project 2 Data Option"
output: pdf_document
date: "2023-03-08"
---

\pagenumbering{gobble}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings=FALSE, message=FALSE)
library(ggplot2)
library(caret)
library(gbm)
library(ggplot2)
library(gridExtra)
library(dplyr)
require(optmatch)
require(MatchIt)
require('gridExtra') 
require(tidyverse)
require('ggplot2')
```

In this project, we try to build a classifier to predict whether the unlabeled test points that are currently identified as "CANDIDIATE" are likely to be a real exoplanet based on other observed characteristics. Our main focus is estimating the uncertainty and accuracy of the classifier.

Initially, we thought of applying conformal inference methods to quantify uncertainty of our classifier. Instead of using residuals as the conformal score, we tried to find a different score function that is appropriate for categorical variables. We define the score as the negative of the logistic regression outcome. This is because the higher the score is, for example the residual, the worse the prediction is, while a larger logit model output means a better predictive power. Therefore choosing negative of the logistic regression outcome would align with the meaning of the scores that we want. However, we soon realized that this was not the best approach because for a binary classification problem, it would not make much sense to predict a set of labels. Typically utilizing conformal inference for classification problems means that instead of a prediction interval, we would get a set of likely labels. For this problem, there are only two possible labels, confirmed or false positive. If we try to get $1-\alpha$ coverage, our interval we may end up with both potential labels which would not tell us much about the accuracy of the classifier.

Another issue with conformal inference methods is that even if they could guarantee a $1-\alpha$ coverage, they require exchangeability of train and test points. Even if the train and test datasets have the same distribution of Y conditional on X, the distributions of X may differ which means the exchangeability assumption would not hold.

# Exploration

We did some initial data cleaning, removing rows with NA values and keeping only the rows that has complete data entries.

We then try two classifiers, logistic regression and gradient boosting. We calculate the accuracy of the two classifiers by using 10 fold cross validation and find that logistic regression has 76% accuracy and gradient boosting has 90% accuracy. We also check for feature importance and see that the variables koi_prad, koi_period, koi_depth and koi_impact are the most important covariates in predicting whether the object is an exoplanet or not. However, it's worth mentioning that these accuracy numbers are actually not a good estimate for the accuracy of the classifier on the unlabeled test set because our train and test set does not have exchangeability and the covariates’ distributions differ. Later we will explore the covariate shift in greater depth by checking their distributions. 

```{r}
koi_df = read.csv("KOI.csv")
analysis_df = koi_df[c("ra","dec","koi_period","koi_time0bk","koi_impact","koi_duration","koi_depth","koi_prad","koi_teq","koi_insol","koi_steff","koi_slogg","koi_srad","koi_disposition")][koi_df$koi_disposition != "CANDIDATE",]
analysis_df$koi_disposition = as.factor(analysis_df$koi_disposition)
analysis_df$koi_disposition <- factor(analysis_df$koi_disposition, labels = c(1,0))
analysis_df = analysis_df[complete.cases(analysis_df), ]
```

```{r}
set.seed(123)
trainIndex <- createDataPartition(analysis_df$koi_disposition, p = .8, 
                                  list = FALSE, 
                                  times = 1)

training <- analysis_df[ trainIndex,]
testing  <- analysis_df[-trainIndex,]

fitControl <- trainControl(## 10-fold CV
                           method = "repeatedcv",
                           number = 10,
                           ## repeated ten times
                           repeats = 1)
```

```{r}
# Fitting gradient boosting takes around 5 minutes
gbmFit1 <- train(koi_disposition ~ ., data = training, 
                 method = "gbm", 
                 trControl = fitControl,
                 verbose = FALSE)
```

```{r}
koi_pred <- predict(gbmFit1, testing)
#confusionMatrix(data = koi_pred, testing$koi_disposition)
gbmImp <- varImp(gbmFit1, scale=FALSE)
#gbmImp
```

```{r, warnings=FALSE, message=FALSE, silent=TRUE, echo=FALSE, results='hide'}
suppressWarnings(logisticFit <- train(koi_disposition ~ ., data = training, 
                method="glm",
                family="binomial",
                trControl = fitControl))
```

```{r}
koi_pred <- predict(logisticFit, testing)
#confusionMatrix(data = koi_pred, testing$koi_disposition)
```

# Checking for covariate shift

As discussed earlier, to use the conformal prediction methods that we learned in class, we need to check for the assumption that exchangeability holds. Specifically, the $X_{n+1}, Y_{n+1}$ from the Unlabeled data set should follow the same distribution as $(X_1,Y_1)...(X_n,Y_n), \space i=1,..n$ from the Labeled data set. If the assumption does not hold, we need to find new ways of determining the accuracy of the classifier on the test set such as assigning weights on the data points based on some criteria.

We attempt to see whether there are significant differences between the distributions of the covariates in the labeled (train) and unlabeled (test) sets. The first thing we try is to fit a classification model across the entire dataset to predict whether a data point is labeled or unlabeled. If the classifier is able to distinguish between labeled and unlabeled points, then it suggests that the two data sets have different distributions and therefore are not exchangeable. 

From running gradient boosting with 10 fold cross validation, we see that the model has 77% accuracy. Since this is much higher than 50%, and the classifier is able to predict quite well whether a data point is labeled or unlabeled, it suggests there is quite a significant difference between the two distributions. Next, we try to find out whether there are specific covariates that have different distributions between the labeled and unlabeled data.

```{r}
# This chunk is directly taken from file "ConformalWeights" file
koi_df = read.csv("KOI.csv")
analysis_df = koi_df[c("ra","dec","koi_period","koi_time0bk","koi_impact","koi_duration","koi_depth","koi_prad","koi_teq","koi_insol","koi_steff","koi_slogg","koi_srad","koi_disposition")]
new_df = koi_df %>%
                  mutate(koi_disposition = recode(koi_disposition,
                                  "CONFIRMED" = "NC",
                                  "FALSE POSITIVE" = "NC",
                                  "CANDIDATE" = "C"))
analysis_df$koi_disposition = as.factor(new_df$koi_disposition)
analysis_df = analysis_df[complete.cases(analysis_df), ]
```

```{r}
# This chunk is directly taken from file "ConformalWeights" file
set.seed(123)
trainIndex <- createDataPartition(analysis_df$koi_disposition, p = .8, 
                                  list = FALSE, 
                                  times = 1)

training <- analysis_df[ trainIndex,]
testing  <- analysis_df[-trainIndex,]

fitControl <- trainControl(## 10-fold CV
                           method = "repeatedcv",
                           number = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary,
                           ## repeated ten times
                           repeats = 1)

gbmFit1 <- train(koi_disposition ~ ., data = training, 
                 metric = "ROC",
                 method = "gbm", 
                 trControl = fitControl,
                 verbose = FALSE)

koi_pred <- predict(gbmFit1, testing)
#confusionMatrix(data = koi_pred, testing$koi_disposition)
``` 

# Determining which covariates shifted

```{r, include=FALSE}
koi_df = read.csv("KOI.csv")

analysis_df = koi_df[c("ra","dec","koi_period","koi_time0bk","koi_impact","koi_duration","koi_depth","koi_prad","koi_teq","koi_insol","koi_steff","koi_slogg","koi_srad","koi_disposition")][koi_df$koi_disposition != "CANDIDATE",]
analysis_df$koi_disposition = as.factor(analysis_df$koi_disposition)
analysis_df$koi_disposition <- factor(analysis_df$koi_disposition, labels = c(1,0))
analysis_df = analysis_df[complete.cases(analysis_df), ]
```

```{r, include=FALSE}
test_df = koi_df[c("ra","dec","koi_period","koi_time0bk","koi_impact","koi_duration","koi_depth","koi_prad","koi_teq","koi_insol","koi_steff","koi_slogg","koi_srad","koi_disposition")][koi_df$koi_disposition == "CANDIDATE",]
test_df$koi_disposition = as.factor(test_df$koi_disposition)
test_df = test_df[complete.cases(test_df), ]
```

There are multiple ways to check for which covariate have a different distribution in the train and test set. Here we include 2 methods: 

1. Visualization including density plots of all covariates that we're interested in, box plot, and scatter plots of pairwise relationship of the four covariates that we consider to be the most important in predicting koi_disposition.  

2. Test statistics. We compared the characteristics such as mean, variance, ecdf etc. of the four covariates to check that there is indeed difference in the distribution.


## Visualization

Here since the range of the independent values are too large with some values as small as around 0, and others as large as around 1000, to make the plot clearer, we apply the logarithmic method on the covariate. For the 9 covariates that we plotted, except for koi_steff and koi_srad, representing the photospheric temperature  and the photospheric radius of the star respectively, seem to show a roughly same distribution, all other covariates are different in their distribution to some extent. Later on we tried selecting four most important covarites (koi_period, koi_impact, koi_depth, koi_prad) with respect to their power in explaining the koi_disposition and tried to match their distributions with the Unlabeled set through methods such as nearest neighbor. And we can see that these four covariates indeed have very different distributions compared with Unlabeled dataset and it's meaningful to do the matching.

```{r, warning=FALSE, echo=FALSE}

p_period <- ggplot() +
  geom_density(data = analysis_df, aes(x = koi_period), fill = "blue", alpha = 0.3) +
  geom_density(data = test_df, aes(x = koi_period), fill = "red", alpha = 0.3) +
  xlab("KOI Period") +
  ylab("Density") +
  scale_x_log10()

p_impact <- ggplot() +
  geom_density(data = analysis_df, aes(x = koi_impact), fill = "blue", alpha = 0.3) +
  geom_density(data = test_df, aes(x = koi_impact), fill = "red", alpha = 0.3) +
  xlab("KOI Impact") +
  ylab("Density") +
  scale_x_log10()

p_duration <- ggplot() +
  geom_density(data = analysis_df, aes(x = koi_duration), fill = "blue", alpha = 0.3) +
  geom_density(data = test_df, aes(x = koi_duration), fill = "red", alpha = 0.3) +
  xlab("KOI Duration") +
  ylab("Density") +
  scale_x_log10()

p_depth <- ggplot() +
  geom_density(data = analysis_df, aes(x = koi_depth), fill = "blue", alpha = 0.3) +
  geom_density(data = test_df, aes(x = koi_depth), fill = "red", alpha = 0.3) +
  xlab("KOI Depth") +
  ylab("Density") +
  scale_x_log10()

p_prad <- ggplot() +
  geom_density(data = analysis_df, aes(x = koi_prad), fill = "blue", alpha = 0.3) +
  geom_density(data = test_df, aes(x = koi_prad), fill = "red", alpha = 0.3) +
  xlab("KOI Prad") +
  ylab("Density") +
  scale_x_log10()

p_teq <- ggplot() +
  geom_density(data = analysis_df, aes(x = koi_teq), fill = "blue", alpha = 0.3) +
  geom_density(data = test_df, aes(x = koi_teq), fill = "red", alpha = 0.3) +
  xlab("KOI Teq") +
  ylab("Density") +
  scale_x_log10()


p_steff <- ggplot() +
  geom_density(data = analysis_df, aes(x = koi_steff), fill = "blue", alpha = 0.3) +
  geom_density(data = test_df, aes(x = koi_steff), fill = "red", alpha = 0.3) +
  xlab("KOI Steff") +
  ylab("Density") +
  scale_x_log10()

p_insol <- ggplot() +
  geom_density(data = analysis_df, aes(x = koi_insol), fill = "blue", alpha = 0.3) +
  geom_density(data = test_df, aes(x = koi_insol), fill = "red", alpha = 0.3) +
  xlab("KOI insol") +
  ylab("Density") +
  scale_x_log10()

p_srad <- ggplot() +
  geom_density(data = analysis_df, aes(x = koi_srad), fill = "blue", alpha = 0.3) +
  geom_density(data = test_df, aes(x = koi_srad), fill = "red", alpha = 0.3) +
  xlab("KOI srad") +
  ylab("Density") +
  scale_x_log10()


grid.arrange(p_period, p_impact, p_duration, p_depth, p_prad, 
             p_teq, p_steff, p_insol, p_srad, ncol = 3, nrow=3)
```

The box plot below also verify that there's difference in distribution of koi_impact between the Labeled and Unlabeled data shown in the mean and variance. Since the information given by the density plot and box plot are pretty similar, here we don't include the rest of the box plot.


```{r, echo=FALSE}
df = rbind(analysis_df, test_df)
df$koi_disposition <- ifelse(df$koi_disposition == 0 | df$koi_disposition == 1, "Label", "Unlabel")

ggplot(df, aes(x = koi_disposition, y = koi_period)) +
  geom_boxplot(aes(fill = koi_disposition)) +
  scale_y_log10()
```

It might also be helpful to check whether the relationship differs between two covariates for the Labeled and Unlabeled data. If the relationship differs greatly, it means that the generalizability of the model is not good and the accuracy of the prediction by the model will be low. Here we looked at a few examples of the relationship between the four important covariates.  

From the outcome we can see that most plots displays similar distributions but there are minor differences. For example, for the koi_period vs koi_disposition plot, we see that there's a clear positive linear relationship on the right end of the x axis (when koi_impact is larger) of the Labeled data, but there is none of the Unlabeled data.

```{r, warning=FALSE, echo=FALSE}
par(mfrow = c(1, 2))

p1 <- ggplot(data=df, aes(x = koi_prad, y = koi_period)) +
  geom_point(aes(color = koi_disposition)) +
  facet_wrap(~koi_disposition) +
  scale_x_log10() +
  scale_y_log10()

p2 <- ggplot(data=df, aes(x = koi_impact, y = koi_period)) +
  geom_point(aes(color = koi_disposition)) +
  facet_wrap(~koi_disposition) +
  scale_x_log10() +
  scale_y_log10()

p3 <- ggplot(data=df, aes(x = koi_depth, y = koi_period)) +
  geom_point(aes(color = koi_disposition)) +
  facet_wrap(~koi_disposition) +
  scale_x_log10() +
  scale_y_log10()

p4 <- ggplot(data=df, aes(x = koi_depth, y = koi_impact)) +
  geom_point(aes(color = koi_disposition)) +
  facet_wrap(~koi_disposition) +
  scale_x_log10() +
  scale_y_log10()

p5 <- ggplot(data=df, aes(x = koi_depth, y = koi_prad)) +
  geom_point(aes(color = koi_disposition)) +
  facet_wrap(~koi_disposition) +
  scale_x_log10() +
  scale_y_log10()

p6 <- ggplot(data=df, aes(x = koi_prad, y = koi_impact)) +
  geom_point(aes(color = koi_disposition)) +
  facet_wrap(~koi_disposition) +
  scale_x_log10() +
  scale_y_log10()

grid.arrange(p1, p2, p3, p4, p5, p6, ncol = 2, nrow=3)
```

```{r, include=FALSE}
#install.packages("MatchIt")
#install.packages("optmatch")
#install.packages("gridExtra")
#require(optmatch)
#require(MatchIt)
#require('gridExtra') 
#require(tidyverse)
#require('ggplot2')
```

```{r, include=FALSE}
data <- read.csv("KOI.csv")
df <- as_tibble(data)
df <- df  %>% filter(!is.na(koi_disposition))
```

```{r, include=FALSE}
require(dplyr)
df = df[c("ra","dec","koi_period","koi_time0bk","koi_impact","koi_duration","koi_depth","koi_prad","koi_teq","koi_insol","koi_steff","koi_slogg","koi_srad","koi_disposition")]
df = df[complete.cases(df), ]
df <- df %>% mutate(train = ifelse(koi_disposition != "CANDIDATE", 0, 1))
df_t <- df
```

Then we check the balance between the labeled and unlabeled data. For the four variables that were the most used in classification during exploratory analysis we find large differences between the datasets, confirming the fact that their distributions are different as we observed in the visualization. The variables we look at are koi_prod, koi_period, koi_impact, and koi_depth. Each variable has imbalance as assessed by difference in means, variance ratio, standardized mean differrence, and eCDF. We summarize the imbalances with a table below. 

```{r}
# Create formula for matching based on train variable and covariates

formula <- train ~ koi_prad + koi_period + koi_impact + koi_depth

# Checking balance 

match_object_before <- matchit(formula, data = df, method = NULL, distance="glm")
summary(match_object_before)
```

# K Nearest Neighbor matching to create a holdout set similar to the test set

We try several different matching techniques. The first is greedy nearest neighbor matching, which involves calculating the distance between every treated unit and control unit. Then, one at a time, each treated unit is paired with a control unit as its match. The matching process is "greedy" because it doesn't optimize. Each match is made without considering any potential matches that may come later in the process. In the table below and in the graph densities we see a much better balance between the covariates, although they are still not ideal.

```{r, }

# Create MatchIt object using nearest neighbor matching
match_object <- matchit(formula, data = df, method = "nearest")

# Summarize the MatchIt object
summary(match_object)
plot(summary(match_object))

# Plot
plot(match_object, type = "jitter", interactive = FALSE)

# Extract matched data
matched_data_nearest <- match.data(match_object)

```


```{r, echo=FALSE, warning=FALSE, }

df <- matched_data_nearest

plot1_after <- ggplot(df, aes(x = koi_prad, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 20) 

plot2_after <- ggplot(df, aes(x = koi_period, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 150) 

plot3_after <- ggplot(df, aes(x = koi_depth, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 2500) 

plot4_after <- ggplot(df, aes(x = koi_impact, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 2.5)
  
grid.arrange(plot1_after, plot2_after, plot3_after, plot4_after,ncol = 1)

```

## Trying alternative approaches to matching

Below we try two improvements to the matching procedure. First, we try an "optimal" procedure that minimizes the sum of the absolute pairwise distances in the matched sample. Then we perform optimal full matching, in which all units, both unlabeled and labeled, receive at least one match. Full matching is an upsampling mthod, which samples with replacement. 

## Optimal

```{r, echo=FALSE, }
#This chunk takes about 5 mins to run
#formula <- train ~ koi_prad + koi_period + koi_impact + koi_depth
df <- df_t 
# Create MatchIt object using nearest neighbor matching
match_object <- matchit(formula, data = df, method = "optimal")

# Summarize the MatchIt object
summary(match_object)
plot(summary(match_object))

# Plot
plot(match_object, type = "jitter", interactive = FALSE)

# Extract matched data
matched_data_optimal <- match.data(match_object)

```

```{r, echo=FALSE, }

df <- matched_data_optimal

plot1_after <- ggplot(df, aes(x = koi_prad, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 20) 

plot2_after <- ggplot(df, aes(x = koi_period, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 150) 

plot3_after <- ggplot(df, aes(x = koi_depth, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 2500) 

plot4_after <- ggplot(df, aes(x = koi_impact, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 2.5)
  
grid.arrange(plot1_after, plot2_after, plot3_after, plot4_after,ncol = 1)

```

## Full

```{r, echo=FALSE,}
# This took around 5 minutes to run

df <- df_t 

#formula <- train ~ koi_prad + koi_period + koi_impact + koi_depth

#plot(summary(match_object_before))

# Create MatchIt object using nearest neighbor matching
match_object <- matchit(formula, data = df, method = "full")

# Summarize the MatchIt object
summary(match_object)
plot(summary(match_object))

# Plot
plot(match_object, type = "jitter", interactive = FALSE)

# Extract matched data
matched_data_full <- match.data(match_object)

```

```{r, echo=FALSE, warning=FALSE, }

df <- matched_data_full

plot1_after <- ggplot(df, aes(x = koi_prad, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 20) 

plot2_after <- ggplot(df, aes(x = koi_period, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 150) 



plot3_after <- ggplot(df, aes(x = koi_depth, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 2500) 


plot4_after <- ggplot(df, aes(x = koi_impact, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 2.5)
  
grid.arrange(plot1_after, plot2_after, plot3_after, plot4_after,ncol = 1)


```

With full matching, we see that often a labeled observation is matched to multiple unlabeled observations:

```{r, echo = FALSE, warning=FALSE, }
ggplot(matched_data_full %>% group_by(subclass) %>% summarize(n = n()) %>% arrange(desc(n)), aes(x=subclass, y=n)) + geom_bar(stat = "identity", fill = "blue")
```

# Using the matched dataset

After using nearest neighbors to create a matched dataset, we check again for exchangeability once more by seeing whether a classifier can accurately predict whether a datapoint is labeled or unlabeled. We see that the accuracy has gone down significantly to 63% comparing to 77%  prior to matching. This suggests that our matched dataset is more likely to have exchangeability with our unlabeled test data.

```{r, include=FALSE}
koi_df = read.csv("KOI.csv")
#koi_df = koi_df[!duplicated(koi_df$kepid), ]
test_set = koi_df[c("ra","dec","koi_period","koi_time0bk","koi_impact","koi_duration","koi_depth","koi_prad","koi_teq","koi_insol","koi_steff","koi_slogg","koi_srad","koi_disposition")][koi_df$koi_disposition == "CANDIDATE",]
test_set = test_set[complete.cases(test_set), ]
```

```{r pressure, echo=FALSE,}
matched_data_nearest = matched_data_nearest[matched_data_nearest$koi_disposition != "CANDIDATE",]
write.csv(matched_data_nearest,"matched_data_nearest.csv")
total = bind_rows(test_set, matched_data_nearest)
total = total[,1:14]
```

```{r, }
# This chunk is directly taken from file "ConformalWeights"
set.seed(123)
new_df = total %>%
                  mutate(koi_disposition = recode(koi_disposition,
                                  "CONFIRMED" = "NC",
                                  "FALSE POSITIVE" = "NC",
                                  "CANDIDATE" = "C"))
total$koi_disposition = as.factor(new_df$koi_disposition)
analysis_df = total[complete.cases(total), ]

trainIndex <- createDataPartition(analysis_df$koi_disposition, p = .8, 
                                  list = FALSE, 
                                  times = 1)

training <- analysis_df[ trainIndex,]
testing  <- analysis_df[-trainIndex,]

fitControl <- trainControl(## 10-fold CV
                           method = "repeatedcv",
                           number = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary,
                           ## repeated ten times
                           repeats = 1)

gbmFit1 <- train(koi_disposition ~ ., data = training, 
                 metric = "ROC",
                 method = "gbm", 
                 trControl = fitControl,
                 verbose = FALSE)

koi_pred <- predict(gbmFit1, testing)
#confusionMatrix(data = koi_pred, testing$koi_disposition)
``` 