---
title: "Nearest Neighbors"
output:
  pdf_document: default
  html_document: default
date: "2023-03-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("MatchIt")
#just a heads up when I ran this it required an update to Rcpp
#install.packages("optmatch")
#install.packages("gridExtra")
require(optmatch)
require(MatchIt)
require('gridExtra') 
require(tidyverse)
require('ggplot2')
```


```{r, include=FALSE}
data <- read.csv("KOI.csv")
df <- as_tibble(data)
df <- df  %>% filter(!is.na(koi_disposition))
```


```{r, include=FALSE}

require(dplyr)
# assuming your data frame is called df

#colSums(is.na(df))

# Added these two lines to keep dataframe consistent with previous analyses
df = df[c("ra","dec","koi_period","koi_time0bk","koi_impact","koi_duration","koi_depth","koi_prad","koi_teq","koi_insol","koi_steff","koi_slogg","koi_srad","koi_disposition")]
df = df[complete.cases(df), ]

# same as above but keeping only 14 variables rather than 40
#df <- df %>% select(-koi_teq_err1, -koi_teq_err2)
# same as above 
df <- df %>% mutate(train = ifelse(koi_disposition != "CANDIDATE", 0, 1))
# same as above
#df <- df %>% filter(!is.na(koi_prad)) %>% filter(!is.na(koi_period)) %>% filter(!is.na(koi_impact)) %>% filter(!is.na(koi_depth)) 
# same as above
df_t <- df 

# df_train <- df %>% filter(koi_disposition != "CANDIDATE")
# df_test <- df %>% filter(koi_disposition == "CANDIDATE")

```


# Demonstrating lack of exchangeability

We first check the balance beetween the labeled and unlabeled data to see if exchangeability holds. For the four variables that were the most used in classification during exploratory analysis we find large differences between the datasets, which suggest a violation of exchangeability. The variables we look at are koi_prod, koi_period, koi_impact, and koi_depth. Each variable has imbalance as assessed by difference in means, variance ratio, standardized mean differrence, and eCDF. We summarize the imbalances with a table below. 

```{r}
# Create formula for matching based on train variable and covariates

formula <- train ~ koi_prad + koi_period + koi_impact + koi_depth

# Checking balance 

match_object_before <- matchit(formula, data = df, method = NULL, distance="glm")
summary(match_object_before)
```

### Remedying the imblance using nearest neighbors

We try several different matching techniques. The first is greedy nearest neighbor matching, which involves calculating the distance between every treated unit and control unit. Then, one at a time, each treated unit is paired with a control unit as its match. The matching process is "greedy" because it doesn't optimize. Each match is made without considering any potential matches that may come later in the process. In the table below and in the graph densities we see a much better balance between the covariates, although they are still not ideal.


```{r}

# Create MatchIt object using nearest neighbor matching
match_object <- matchit(formula, data = df, method = "nearest")

# Summarize the MatchIt object
summary(match_object)
plot(summary(match_object))

# Plot
plot(match_object, type = "jitter", interactive = FALSE)

# Extract matched data
matched_data_nearest <- match.data(match_object)

```
```{r, echo=FALSE}

df <- matched_data_nearest

plot1_after <- ggplot(df, aes(x = koi_prad, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 20) 

plot2_after <- ggplot(df, aes(x = koi_period, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 150) 

plot3_after <- ggplot(df, aes(x = koi_depth, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 2500) 

plot4_after <- ggplot(df, aes(x = koi_impact, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 2.5)
  
grid.arrange(plot1_after, plot2_after, plot3_after, plot4_after,ncol = 1)

```

### Trying alternative approaches to matching

Below we try two improvements to the matching procedure. First, we try an "optimal" procedure that minimizes the sum of the absolute pairwise distances in teh matched sample. Then we perform optimal full matching, in which all units, both unlabeled and labeled, receive at least one match. Full matching is an upsampling mthod, which samples with replacement. 

## Optimal


```{r, echo=FALSE}
#This chunk takes about 5 mins to run
#formula <- train ~ koi_prad + koi_period + koi_impact + koi_depth
#df <- df_t 
# Create MatchIt object using nearest neighbor matching
match_object <- matchit(formula, data = df, method = "optimal")

# Summarize the MatchIt object
summary(match_object)
plot(summary(match_object))

# Plot
plot(match_object, type = "jitter", interactive = FALSE)

# Extract matched data
matched_data_optimal <- match.data(match_object)

```

```{r, echo=FALSE}

df <- matched_data_optimal

plot1_after <- ggplot(df, aes(x = koi_prad, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 20) 

plot2_after <- ggplot(df, aes(x = koi_period, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 150) 

plot3_after <- ggplot(df, aes(x = koi_depth, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 2500) 

plot4_after <- ggplot(df, aes(x = koi_impact, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 2.5)
  
grid.arrange(plot1_after, plot2_after, plot3_after, plot4_after,ncol = 1)

```

## Full

```{r, echo=FALSE}
# This took around 5 minutes to run

#df <- df_t 

#formula <- train ~ koi_prad + koi_period + koi_impact + koi_depth

#plot(summary(match_object_before))

# Create MatchIt object using nearest neighbor matching
match_object <- matchit(formula, data = df, method = "full")

# Summarize the MatchIt object
summary(match_object)
plot(summary(match_object))

# Plot
plot(match_object, type = "jitter", interactive = FALSE)

# Extract matched data
matched_data_full <- match.data(match_object)

```
```{r, echo=FALSE}

df <- matched_data_full

plot1_after <- ggplot(df, aes(x = koi_prad, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 20) 

plot2_after <- ggplot(df, aes(x = koi_period, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 150) 



plot3_after <- ggplot(df, aes(x = koi_depth, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 2500) 


plot4_after <- ggplot(df, aes(x = koi_impact, fill= factor(train))) +  geom_density(position="identity", alpha = 0.5) + xlim(0, 2.5)
  
grid.arrange(plot1_after, plot2_after, plot3_after, plot4_after,ncol = 1)


```

With full matching, we see that often a labeled observation is matched to multiple unlabeled observations:

```{r, echo = FALSE}

ggplot(matched_data_full %>% group_by(subclass) %>% summarize(n = n()) %>% arrange(desc(n)), aes(x=subclass, y=n)) + geom_bar(stat = "identity", fill = "blue")

```


